Dynamic Space Efficient Cuckoo Table when doubling is too much -- a novel approach to growing!

### 2017-04-11 paper/paper.tex
Laptop - Paper Stuff

\subjclass{Dummy classification -- please refer to \url{http://www.acm.org/about/class/ccs98-html}}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation".
\keywords{Dynamic data-structures, Open addressing, Closed hashing}% mandatory: Please provide 1-5 keywords
% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
A lot of previous work concentrates on optimizing the speed and space
efficiency of hash tables.  Additionally many practical
implementations exist.  Modern displacement strategies can fill a
given table to more than 95\% before the performance begins to decrease
significantly.  But once the threshold is met, copying and migrating
the whole table still seems to be the most efficient answer.

To achieve any kind of efficiency, each such migration must be
amortized by newly inserted elements.  Therefore, each migration has
to increase the capacity by a significant amount -- usually, this
means doubling the capacity.  Growing by a constant factor
$\gamma$ often leads to bad space efficiency because the newly grown table
is only filled to less than a factor of $\gamma^{-1}$.

The space efficiency is even worse during the migration, where both
the new and the old table coexist at the same time causing a massive
space overhead.  To solve both of these issues, we propose the a new
hash table architecture that supports dynamic growing while keeping a
strict size constraint $\alpha \cdot n$ throughout its lifetime.  Even
under these conditions, we can guarantee constant lookup times, and
amortized insertion times, independent of $n$.
 \end{abstract}

@@

\section{Preliminaries}
A hash table is a data-structure, which stores key-value-pairs
(\(\langle key, data \rangle\)) and offers the following functions:
\verb~insert~ stores a given key-value pair, \verb~find~ given a key returns if
said key was stored, and if so it returns the stored value, and
\verb~erase~ which deletes a previously inserted element (if present).

@@

For large elements (\(\gg size(pointer)\)), one can simply use open
hashing, to reduce the relevant memory factor.  Therefore, we restrict
ourselves to the common case of elements whose size is comparable to
that of a pointer.  For our experiments we use 128bit elements (64bit
keys and 64bit values).  In these instances closed addressing
introduces a significant memory overhead (at least \(\times1.5\)),
therefore, we only consider closed hash tables for the purpose of this
paper. For closed hash tables, the memory efficiency is directly
dependent on the element density.

\subsection{$\alpha$-Space Efficient Hash Tables}
We call a hashing technique \(\alpha\)-space efficient when it can
work effectively using at most $\alpha \cdot size_{\min} + O(1)$
memory (TODO DEFINE WORKING EFFICIENTLY FURTHER i.e. constant average
running time or just look at experiments or inserting the last 10\%
elements is less than inserting all others). In this case we define
working efficiently as having average insertion times close to
$O(\frac{m}{m-n})$ (average number of probes to hit an empty cell).

@@

\subsubsection*{$\alpha$-Space Efficiency for Dynamic Tables}
The definition of a space efficient hashing technique given in (TODO
reference) the previous section is specifically targeted for
statically sized hash tables.  As detailed in Motivation (TODO
REFERENCE) our goal is to construct a hash table which can efficiently
(and tightly) grow with the number of inserted elements. Therefore, we
want to define the notion of an \(\alpha\)-space efficient dynamic
table.

We call a hash table implementation dynamically \(\alpha\)-space
efficient if an instantiated table can grow arbitrarily large over its
original capacity, while remaining smaller than \(\alpha\cdot
mem_{necessary}\) at all times.  \[mem_{curr} \leq \alpha\cdot
max(n_{curr})\cdot size(element)\]

One point of discussion while defining dynamic space efficiency is the
space consumption during hash table operations.  During a table
migration both the source (\(t_{source}\)) and the target
(\(t_{target}\)) table are allocated, thus using \(mem =
(m_{source}+m_{target})\cdot size(cell)\) space.  This is especially
prohibiting if all elements are held in one table, then at least
\(2\cdot m\cdot size(element)\) is necessary for each migration.

@@

\paragraph{Trivial Examples}
One simple example of this is a linear probing table, which grows by a
factor of two, whenever 50\% of the table is filled. Here the table is
at least 25\% filled (immediately after growing), therefore, the table
is 6-space efficient (during the grow operation there are \((4+2)+n\)
cells). Another example would be a cuckoo table which doubles in size
whenever it surpasses 95\%, fill rate. This would be a
$0.95^{-1}+0.45^{-1}\approx 3.2$-space efficient table.

\subsection{Cuckoo Hashing}
Cuckoo hashing is a technique to resolve hash conflicts in an closed hashing hash table. Its main draw is that it guarantees constant
lookup times even in densely filled tables. The article TODO by TODO
(probably mitzenmacher with or without dietzfelbinger) et al. gives a
good overview over many cuckoo hashing variants.  The distinguishing
technique of cuckoo hashing is, that $H$ hash functions (\(h_1, ... ,
h_H\)) are used to find $H$ independent possible positions. Each
element is stored in one of its positions.  Even if all possible
positions are occupied one can often move elements to create space for the
current element. We call this displacing elements.

@@

\section{DySECT Space Efficient Growing, by Splitting the Table}
Our goal is to build an efficiently growing data structure, that
remains space efficient at all times.  A commonly used growing
technique is to double the previous size by migrating all elements
into a table which has two times the original size.  Of course this is
not memory efficient.  A natural thought which occurs is to double
only part of the data structure.  This idea together with the
balancing ability of multiple hash functions is what will achieve the
described functionality.

\subsection{Description}
We use a two level approach (shown in TODO FIGREFRENCE) to allow
differentiated growing between table parts.  The first level consists
of a table which stores pointers to $T$ second level tables. The
second level tables consist of buckets that store elements.

Comparable to classic cuckoo hashing, each element has $H$ associated
buckets ($H$ number of hash functions), these can be in the same, or
in different sub tables.  Each inserted element lies in one of its
associated buckets.  To find a bucket associated with an
element $e$, we compute \(e\)s hash value using the appropriated hash function
$h_i(e)$. We split that hash value into two parts, then we use the
first part to chose a second level table, and the second part to
choose one of the contained buckets.

To understand the performance of this hash table, it is important to
realize that when the hash table is used regularly, the first level
table will always remain cached. This is important, because otherwise lookups
would cause unnecessary cache misses that would not have happened in a
single level hash table.  Because of this we expect the average number
of cache misses during hash table operations to be similar to those of
a normal cuckoo hash table.

@@

Assume that all subtables have $s=m/tl$ cells. We can grow the first
subtable once $\alpha\cdot n > m+2s$.  Doubling the size of a second
level table increases the global number of cells from $m_{old} =
tl\cdot s$ to $m_{new} = m_{old}+s = (tl+1)\cdot s$ (factor:
$\frac{tl+1}{tl}$).  Note that all subsequent grows migrate one of the
smaller tables, until all tables have the same size again.  Therefore,
each grow until then increase the capacity by the same absolute amount
(smaller factor).

@@

Shrinking can work similarly to growing. We replace a second level
table with a smaller one by migrating elements from one to the
other.  During this migration we join elements from two buckets into
one. Therefore, it is possible for a bucket to overfill.  We reinsert
these elements at the end of the migration.  Obviously, this can only
apply to at most half the migrated elements.

@@

\subsubsection*{Reducing Number Computed Hash Functions}
Evaluating hash functions is expensive, therefore, reducing the number
of hash functions computed per operation increases the efficiency of
the table.  The hash function we use computes 64bit hash values
(i.e. xxHash TODO).  We split these 64bit value into two 32bit hash
values. All common bucket hash table sizes can be addressed using 32
bits. They can address up to $2^{32}$ buckets which achieves a
capacity for \(2^{35} \approx 34\) billion elements (bucket size 8),
at that point a hash table consumes 512GiB memory.

@@

\subsubsection*{Imbalance through Inhomogeneous Table Resolution}
\label{sec:inhom_res}
By growing individual second level tables we introduce a size imbalance
between subtables.  Large tables contain more buckets, but the number
of elements hashed to the table itself is not dependent on its size,
therefore, it is difficult to spread elements evenly among buckets.
Uneven bucket fill ratios can lead to longer insertion times.

If there are $n$ elements in a table with \emph{tl} second level tables,
$j$ of which have size $2s$ the others have size $s$. If elements are
spread equally among buckets then all small tables have around
$n/(tl+j)$ elements, and the bigger tables have $2n/(tl+j)$ elements.
For each table there are about $kn/tl$ elements associated to one of
its buckets.  This shows that having more hash functions can lead to a
better balance.

For two hash functions ($k=2$) and only one grown table ($j=1$) this
means that $\approx 2n/(tl+1)$ should be stored in the first table.
These are nearly all elements associated to a bucket in the first
table ($\approx 2n/tl$). So to distribute elements evenly nearly all
elements would have to be stored there.

@@

\subsection{Comparison Implementations}
To put the performance of our solution into perspective, we
implemented and tested several other options for space efficient
hashing.  We use our own implementations, since most hash tables found
online do not support strict space-efficiency.  In Appendix TODO we
show that our implementations are at least competitive with
implementations found .

\textbf{Linear Probing} is the most common hashing technique outside of
hashing with chaining.  It is commonly accepted to be inefficient for
higher load factors because search distances can become large.  In
practice, this fact is counteracted by its cache-efficiency making
operations fast in spite of long probing distances.  Linear probing is
especially slow in \verb~find~ and \verb~erase~ operations, that often have
guaranteed constant running time in other implementation.

@@

\textbf{Hopscotch Hashing} is a technique similar to linear probing where
elements are moved when they are too far from their original hash
position.  Additionally, hopscotch hashing introduces an acceleration
data structure which uses bitmaps to store neighborhood data. This
data improves find and erase performance.  Many people would expect
the acceleration data to be the main draw to hopscotch hashing, but it
is also its greatest demise in the space-efficient setting.  The
bitmaps introduce a significant memory overhead of neighborhood-size
bits per cell.  Our experiments show that for any reasonable fill
factor neighborhoods have to be 64bits and larger, this introduces a
size factor of $\times1.5$ (128bit elements).  We still include
measurements with hopscotch hashing -- without considering space for
the neighborhood data -- to show its overall performance tendencies.
Growing hopscotch tables is done similar to the technique described
under linear probing.

\textbf{Robin Hood Hashing} is another technique based on linear probing. The
goal of this technique is to improve find performance by minimizing
the maximum/average displacement distance.  The drawback of this technique is
that insertions have to compute the hash of each element that is
skipped by the insertion.

\textbf{Classic Bucket Cuckoo Hashing} is the technique we adopted to design
our hashing technique.  It has clear advantages for space efficient
tables.  Like our variant it offers constant lookup and deletion
times.  In contrast to our variant, it does not suffer from any
imbalances caused either by table resolution or by local growing
operations.  Growing works similar to the linear probing table above.

\subsection{Hardware}
We test our implementation on two very different systems, first a
consumer level desktop computer \ldots{}

\subsection{Tests with a Static Table Size}
Our hash table is designed for space efficient growing, therefore, its
main advantages show in dynamic workloads.  We still performed some
tests with a static table size.  With these measurements, we can
highlight different points without the influence of dynamically growing.

@@

DySECT fails earlier than cuckoo hashing.  The cause for this is
twofold. First, our architecture ($T$ subtables with power of two
sizes) does not allow exactly $24M$ cells, therefore, our table is
slightly more full. Secondly, to get close to $24M$ cells, we have
$110$ large subtables (twice the size of the others).  This introduces
inhomogeneous table resolution (see \ref{sec:inhom_res}), which
decreases our performance compared to cuckoo hashing.  This test shows
the influence of inhomogeneous table resolution without the influence
of size changes, and the connected table imbalance.

@@

\subsubsection*{Constant Sized Table Under Varying Fill Degree}
\begin{figure}
  \includegraphics[width=\textwidth]{../eval/plots/eps_insert.pdf}
  \caption{Insert timings at different fill degrees into a non-growing table.}
  \label{fig:eps_insert}
\end{figure}
\begin{figure}
  \includegraphics[width=0.48\textwidth]{../eval/plots/eps_find_s.pdf}
  \includegraphics[width=0.48\textwidth]{../eval/plots/eps_find_u.pdf}
  \caption{Timings for find operations with different fill degrees (\emph{left:}successful, \emph{right:}unsuccessful).}
  \label{fig:eps_find}
\end{figure}

\subsubsection*{Constructing a Table with Growing}

\begin{figure}
  \includegraphics[width=\textwidth]{../eval/plots/ti_insert.pdf}
  \caption{Average time per insert operation during the construction of a dynamic table.}
  \label{fig:ti_insert}
\end{figure}
\begin{figure}
  \includegraphics[width=0.48\textwidth]{../eval/plots/ti_find_s.pdf}
  \includegraphics[width=0.48\textwidth]{../eval/plots/ti_find_u.pdf}
  \caption{Timings for find operations on the grown table (from Figure~\ref{fig:ti_insert}; \emph{left:}successful, \emph{right:}unsuccessful).}
  \label{fig:ti_find}
\end{figure}

>>>>

\subjclass{Dummy classification -- please refer to \url{http://www.acm.org/about/class/ccs98-html}}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation".
\keywords{Dynamic data-structures, Open addressing, Closed hashing, Cuckoo hashing, Dense hash tables}% mandatory: Please provide 1-5 keywords
% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
A lot of previous work concentrates on optimizing the speed and space
efficiency of hash tables.  Additionally many practical
implementations exist.  Modern open addressing techniques can fill a
given table to more than 95\% before the performance deteriorates
significantly.  But once the threshold is met, copying and migrating
the whole table still seems to be the most efficient answer.

To achieve any kind of efficiency, each such migration must be
amortized by newly inserted elements.  Therefore, each migration has
to increase the capacity by a significant amount -- usually, this
means doubling the capacity.  Growing by large factor
$\gamma$ often leads to bad space efficiency because the newly grown table
is only filled to less than a factor of $\gamma^{-1}$.

The space efficiency is even worse during the migration, where
oftentimes both the new and the old table coexist at the same time
causing a massive space overhead.  To solve both of these issues, we
propose the a new hash table architecture that supports dynamic
growing while keeping a strict size constraint $\alpha \cdot n$
throughout its lifetime.  Even under these conditions, we can
guarantee constant lookup times, and amortized insertion times,
independent of $n$.
 \end{abstract}

@@

\section{Preliminaries}
A hash table is a data-structure, which stores key-value-pairs
($\langle key, data \rangle$) and offers the following functions:
\verb~insert~ stores a given key-value pair, \verb~find~ given a key returns if
said key was stored, and if so it returns the stored value, and
\verb~erase~ which deletes a previously inserted element (if present).

@@

For large elements ($\gg size(pointer)$), one can simply use open
hashing, to reduce the relevant memory factor.  Therefore, we restrict
ourselves to the common case of elements whose size is comparable to
that of a pointer.  For our experiments we use 128bit elements (64bit
keys and 64bit values).  In these instances closed addressing
introduces a significant memory overhead (at least $\times1.5$),
therefore, we only consider closed hash tables for the purpose of this
paper. For closed hash tables, the memory efficiency is directly
dependent on the element density.

\subsection{$\alpha$-Space Efficient Hash Tables}
We call a hashing technique $\alpha$-space efficient when it can
work effectively using at most $\alpha \cdot size_{\min} + O(1)$
memory (TODO DEFINE WORKING EFFICIENTLY FURTHER i.e. constant average
running time or just look at experiments or inserting the last 10\%
elements is less than inserting all others). In this case we define
working efficiently as having average insertion times close to
$O(\frac{m}{m-n})$ (average number of probes to hit an empty cell).

@@

\subsubsection*{$\alpha$-Space Efficiency for Dynamic Tables}
The definition of a space efficient hashing technique given in (TODO
reference) the previous section is specifically targeted for
statically sized hash tables.  As detailed in Motivation (TODO
REFERENCE) our goal is to construct a hash table which can efficiently
(and tightly) grow with the number of inserted elements. Therefore, we
want to define the notion of an $\alpha$-space efficient dynamic
table.

We call a hash table implementation dynamically $\alpha$-space
efficient if an instantiated table can grow arbitrarily large over its
original capacity, while remaining smaller than $\alpha\cdot
mem_{necessary}$ at all times.
\[mem_{curr} \leq \alpha\cdot max(n_{curr})\cdot size(element)\]

One point of discussion while defining dynamic space efficiency is the
space consumption during hash table operations.  During a table
migration both the source ($t_{source}$) and the target
($t_{target}$) table are allocated, thus using $mem =
(m_{source}+m_{target})\cdot size(cell)$ space.  This is especially
prohibiting if all elements are held in one table, then at least
$2\cdot m\cdot size(element)$ is necessary for each migration.

@@

\paragraph*{Trivial Examples}
One simple example of this is a linear probing table, which grows by a
factor of two, whenever 50\% of the table is filled. Here the table is
at least 25\% filled (immediately after growing), therefore, the table
is 6-space efficient (during the grow operation there are $(4+2)+n$
cells). Another example would be a cuckoo table which doubles in size
whenever it surpasses 95\%, fill rate. This would be a
$0.95^{-1}+0.45^{-1}\approx 3.2$-space efficient table.

\subsection{Cuckoo Hashing}
Cuckoo hashing is a technique to resolve hash conflicts in an closed
hashing hash table. Its main draw is that it guarantees constant
lookup times even in densely filled tables. The article TODO by TODO
(probably mitzenmacher with or without dietzfelbinger) et al. gives a
good overview over many cuckoo hashing variants.  The distinguishing
technique of cuckoo hashing is, that $H$ hash functions ($h_1, ... ,
h_H$) are used to find $H$ independent possible positions. Each
element is stored in one of its positions.  Even if all possible
positions are occupied one can often move elements to create space for
the current element. We call this displacing elements.

@@

\section{DySECT Space Efficient Growing, by Splitting the Table}
Our goal is to build an efficiently growing data structure, that
remains space efficient at all times.  A commonly used growing
technique is to increase the size by migrating all elements
into a larger new table.  Of course this is
not memory efficient.  A natural thought which occurs is to double
only part of the data structure.  This idea together with the
balancing ability of multiple hash functions is what will achieve the
described functionality.

\subsection{Description}
We use multiple ($T$) subtables (shown in TODO FIGREFRENCE) to allow
differentiated growing between table parts.  The subtables consist of
buckets that store elements.

Comparable to classic cuckoo hashing, each element has $H$ associated
buckets ($H$ number of hash functions), these can be in the same, or
in different subtables.  Each inserted element lies in one of its
associated buckets.  To find a bucket associated with an
element $e$, we compute $e$s hash value using the appropriated hash function
$h_i(e)$. We split that hash value into two parts, then we use the
first part to chose a subtable, and the second part to
choose one of the contained buckets.

%% To understand the performance of this hash table, it is important to
%% realize that when the hash table is used regularly, the first level
%% table will always remain cached. This is important, because otherwise lookups
%% would cause unnecessary cache misses that would not have happened in a
%% single level hash table.  Because of this we expect the average number
%% of cache misses during hash table operations to be similar to those of
%% a normal cuckoo hash table.

@@

Assume that all subtables have $s=m/T$ cells. When $\alpha\cdot n >
m+2s$ We can grow the first subtable while keeping the size
constraint.  Doubling the size of a subtable increases the global
number of cells from $m_{old} = T\cdot s$ to $m_{new} = m_{old}+s =
(T+1)\cdot s$ (factor: $\frac{T+1}{T}$).  Note that all subsequent
grows migrate one of the smaller tables, until all tables have the
same size again.  Therefore, each grow until then increase the
capacity by the same absolute amount (smaller factor).

@@

Shrinking can work similarly to growing. We replace a subtable with a
smaller one by migrating elements from one to the other.  During this
migration we join elements from two buckets into one. Therefore, it is
possible for a bucket to overfill.  We reinsert these elements at the
end of the migration.  Obviously, this can only apply to at most half
the migrated elements.

@@

\subsubsection*{Reducing Number Computed Hash Functions}
Evaluating hash functions is expensive, therefore, reducing the number
of hash functions computed per operation increases the efficiency of
the table.  The hash function we use computes 64bit hash values
(i.e. xxHash TODO).  We split these 64bit value into two 32bit hash
values. All common bucket hash table sizes can be addressed using 32
bits. They can address up to $2^{32}$ buckets which achieves a
capacity for $2^{35} \approx 34$ billion elements (bucket size 8),
at that point a hash table consumes 512GiB memory.

@@

\subsubsection*{Imbalance through Inhomogeneous Table Resolution}
\label{sec:inhom_res}
By growing subtables individualy we introduce a size imbalance
between subtables.  Large tables contain more buckets, but the number
of elements hashed to the table itself is not dependent on its size,
therefore, it is difficult to spread elements evenly among buckets.
Uneven bucket fill ratios can lead to longer insertion times.

If there are $n$ elements in a hash table with $T$ subtables,
$j$ of which have size $2s$ the others have size $s$. If elements are
spread equally among buckets then all small tables have around
$n/(T+j)$ elements, and the bigger tables have $2n/(T+j)$ elements.
For each table there are about $Hn/T$ elements associated to one of
its buckets.  This shows that having more hash functions can lead to a
better balance.

For two hash functions ($H=2$) and only one grown table ($j=1$) this
means that $\approx 2n/(T+1)$ should be stored in the first table.
These are nearly all elements associated to a bucket in the first
table ($\approx 2n/T$). So to distribute elements evenly nearly all
elements would have to be stored there.

@@

\subsection{Methodology and Comparison Implementations}

\subsubsection*{Choice of Constants}
In our tests, we use three hash functions ($H=3$) and a bucket size of
($B=8$). These values have consistently outperformed other
possibilities.  We chose to use 256 subtables for all our tests
($T=256$).  The number of subtables only plays a minor role for the
running time.

\subsubsection*{Comparison Implementations}
To put the performance of our solution into perspective, we
implemented and tested several other options for space efficient
hashing.  We use our own implementations, since most hash tables found
online do not support strict space-efficiency.  In Appendix TODO we
show that our implementations are at least competitive with
implementations found .

\paragraph*{Linear Probing} is the most common hashing technique outside of
hashing with chaining.  It is commonly accepted to be inefficient for
higher load factors because search distances can become large.  In
practice, this fact is counteracted by its cache-efficiency making
operations fast in spite of long probing distances.  Linear probing is
especially slow in \verb~find~ and \verb~erase~ operations, that often have
guaranteed constant running time in other implementation.

@@

\paragraph*{Hopscotch Hashing} is a technique similar to linear probing where
elements are moved when they are too far from their original hash
position.  Additionally, hopscotch hashing introduces an acceleration
data structure which uses bitmaps to store neighborhood data. This
data improves find and erase performance.  Many people would expect
the acceleration data to be the main draw to hopscotch hashing, but it
is also its greatest demise in the space-efficient setting.  The
bitmaps introduce a significant memory overhead of neighborhood-size
bits per cell.  Our experiments show that for any reasonable fill
factor neighborhoods have to be 64bits and larger, this introduces a
size factor of $\times1.5$ (128bit elements).  We still include
measurements with hopscotch hashing -- without considering space for
the neighborhood data -- to show its overall performance tendencies.
Growing hopscotch tables is done similar to the technique described
under linear probing.

\paragraph*{Robin Hood Hashing} is another technique based on linear probing. The
goal of this technique is to improve find performance by minimizing
the maximum/average displacement distance.  The drawback of this technique is
that insertions have to compute the hash of each element that is
skipped by the insertion.

\paragraph*{Classic Bucket Cuckoo Hashing} is the technique we adopted to
design our hashing technique.  It has clear advantages for space
efficient tables.  Like our variant it offers constant lookup and
deletion times.  In contrast to our variant, it does not suffer from
any imbalances caused either by table resolution or by local growing
operations.  Growing works similar to the linear probing table above.
Similar to DySECT, $H=3$ and $B=8$ have turned out to be an optimum
for performance.

\subsubsection*{Hardware}
We test our implementation on two very different systems, first a
consumer level desktop computer \ldots{}

\subsection{Influence of Fill Ratio under a Static Table Size}
\label{sec:exp_eps}
\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{../eval/plots/eps_insert.pdf}
  \caption{\label{fig:eps_insert}Insert timings at different fill
    degrees into a non-growing table.}
\end{figure}
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.49\textwidth]{../eval/plots/eps_find_s.pdf}
  \includegraphics[width=0.49\textwidth]{../eval/plots/eps_find_u.pdf}
  \caption{Timings for find operations with different fill degrees
    (\emph{left:}successful, \emph{right:}unsuccessful).}
  \label{fig:eps_find}
\end{figure}
Our hash table is designed for space efficient growing, therefore, its
main advantages show in dynamic workloads.  We still performed some
tests with a static table size.  With these measurements, we can
highlight different points without the influence of dynamically growing.

@@

DySECTs performance degrades earlier than cuckoo hashing.  The cause
for this is twofold. First, our architecture -- $T$ subtables with power
of two sizes -- does not allow exactly
$24M$ cells , therefore, our table is slightly more full. Secondly, to
get close to $24M$ cells, we have $110$ large subtables (twice the
size of the others).  This introduces inhomogeneous table resolution
(see \ref{sec:inhom_res}), which decreases our performance compared to
cuckoo hashing.  This test shows the influence of inhomogeneous table
resolution without the influence of size changes, and the connected
table imbalance.

@@

\subsection{Constructing a Table with Growing}
\label{sec:exp_ti}
\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{../eval/plots/ti_insert.pdf}
  \caption{\label{fig:ti_insert}Average time per insert operation
    during the construction of a dynamic table.}
\end{figure}
%% \begin{figure}
%%   \includegraphics[width=0.48\textwidth]{../eval/plots/ti_find_s.pdf}
%%   \includegraphics[width=0.48\textwidth]{../eval/plots/ti_find_u.pdf}
%%   \caption{Timings for find operations on the grown table (from Figure~\ref{fig:ti_insert}; \emph{left:}successful, \emph{right:}unsuccessful).}
%%   \label{fig:ti_find}
%% \end{figure}
In this test $20M$ elements are inserted into a previously empty
table. All tables are initialized expecting 50\,000 elements, thus
growing is necessary, to fit all elements.  The tables are configured,
to guarantee a load factor of
$\varepsilon$. Figure~\ref{fig:ti_insert} shows the performance in
relation to $\varepsilon$.  Insertion times are normalized similar to
Figure~\ref{fig:eps_insert} (divided by $\frac{1}{\varepsilon}$) to
make them more readable.

We see that DySECT performs by far the best even with less filled
tables.  Here we achieve a speedup of TODO$\%$ over the next best
solution (blabla vs. linear probing blabla at $\varepsilon = 0.$ bla).
On denser instances, we can increase this margin to TODO$\%$ (blabla
vs we'll see blabla at $\varepsilon = 0.$ bla).  With shrinking
$\varepsilon$, we see that insertion times start to degrade for our
competitors.  DySECT remains close to $O(\frac{1}/{\varepsilon})$ even
for fill degrees up to 97.5$\%$ (TODO check).

We also measured the performance of find operations on the created
tables, they are similar to the performance on the static table in
Section~\ref{sec:exp_eps} (see Figure~\ref{fig:eps_find}), therefore,
we omit plotting them for space reasons.

\subsection{Word Count a Practical use Case}
One of the most common use cases for hash tables -- dynamic hash
tables in particular -- is to unify data elements with the same key.
Oftentimes it is unclear how many unique keys are present (how large
the table will get).  In these cases, only dynamic hash tables can
remain space efficient.  In the following test, we take the first
block of the common crawl data-set (TODO REFERENCE) and compute a word
count of the contained words.  The chosen block contains around 240M
words, with around 20M unique words.

For the test, we hash each word and insert it together with a counter.
Subsequent insertions of a reocurring word increase this counter.
Similar to the growing benchmark, we start with an empty table
initialized for 50\,000 elements.  The performance results can be seen
in Figure~\ref{fig:crawl}.  We do not use any normalization since --
on average -- each word is repeated 12 times, therefore, most inserts
will actually behave like successful find operations (returning an
iterator to the contained element).  Our DySECT table manages to
dominate the field, similar to the growing test in
Section~\ref{sec:exp_ti}.  This is a hint to the fact, that insert
performance is important even in find intensive workloads.  This
insight is intensified in the following test
(Section~\ref{sec:exp_mix}).

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{../eval/plots/crawl.pdf}
  \caption{\label{fig:crawl} }
\end{figure}

\subsection{Combining insert and find operations}
In this test, we simulate an application in progress.  We use tables
with a minimum load factor of $\varepsilon = 0.9xx$ (TODO choose one).
The test starts a decently filled table (TODO M elements).  On this
table we perform a mixture of insert and find operations.  Even on
